2D Subquadratic Separable Distance Transformation for Path-Based Norms\thanks{This work has been mainly funded by ANR-11-BS02-009 and ANR-11-IDEX-0007-02 PALSE/2013/21 research grants.}

### 2014-02-21 latex/dgci2014/article.tex
release

Since early works on digital geometry, distance transformation plays
an important role in many applications
\cite{Rosenfeld1966,Rosenfeld1968}. Given a finite input shape
$X\subset \Z^n$, the distance transformation labels each point in $X$
with the distance to its closest point in $\Z^n \setminus X$ (labeling
each point by the closest background point leads to Voronoi maps, see
below). Since such characterization is parametrized by a distance
function, many works have been done to address this distance
transformation problem with a trade-off between algorithmic
performances and the \emph{accuracy} of the digital distance function
with respect to the Euclidean one.  Hence, authors have considered
distances based on chamfer masks
\cite{Rosenfeld1968,borgefors,fouard:ivc:2005} or sequences of chamfer
masks \cite{ROSEN_66,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13};
the vector displacement based Euclidean distance
\cite{danielson,ragnemalm}; the Voronoi diagram based Euclidean
distance \cite{BreuEtAl95,maurer_pami} or the square of the Euclidean
distance \cite{Hirata,roerdnik}.  For Euclidean or $L_p$ metrics,
separable volumetric computations have demonstrated to be very
efficient: optimal $O(n\cdot N^n)$ time algorithms for shapes in $N^n$
domains, optimal multi-thread/GPU implementation\ldots (please refer
to \cite{dcoeurjo_ChapDTWADGMM} for a discussion). For path-based
metrics (chamfer mask, -weighted- neighborhood sequences,\ldots), two
main techniques exist to compute the distance transformation. The
first one considers a weighted graph formulation of the problem, and
shortest path in graphs using Dijkstra-like algorithm can be
considered. If $m$ denotes the size of the chamfer mask, computational
cost could be in $O(m\cdot N^n)$ using a cyclic bucket data structure
\cite{verwer_uniform}. Another approach consists in a raster scan of
the domain: first the chamfer mask is decomposed into two disjoint
sub-masks. Then the domain grid points are scanned in a given order
(consistent with the sub-masks construction) and a local minimum
computation is performed and then propagated
\cite{ROSEN_66,borgefors}. Scanning the domain twice for each
sub-masks leads to the distance transformation. Again, we end up with
a $O(m\cdot N^n)$ computational cost. This complexity is said to be
quadratic in $m$ and $N$ since for arbitrary large image $N$, even if
$m \ll N$, $m$ has to be in $O(N^n)$ to have accurate asymptotic DT.
Please note that Dijkstra's graph approach allows us to defined
constrained distance transformation (\emph{i.e.} geodesic metric),
both separable approaches and raster-scan for path-based metrics are
only dedicated to compact convex domains (usually hyper-rectangular)
distance transformation.

\textbf{Contributions} In this article, we first describe the generic
framework for separable distance transformation and conditions on the
metric to be consistent with this model. Then, we describe
subquadratic algorithms in dimension 2 to compute error-free distance
transformation and Voronoi map for chamfer norms and other path-based
metrics. Overall computational costs can be summarized as follows:
%\begin{table}
%  \caption{Computational cost summary for separable Voronoi map
%    computation on $N^n$ domains ($m$ being the size of the chamfer
%    norm and $f$ the number of row in a H-representation, see below).}
 % \label{tab:final}
  \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      Metric &\textsc{Closest}& \textsc{HiddenBy} & Sep. Voronoi Map & Reference\\
      \hline
      $L_2$ & $O(1)$ &  $O(1)$ & $\Theta(n\cdot N^n)$ & \cite{Hirata1996}\\
      $L_\infty$ & $O(1)$ & $O(1)$ &  $\Theta(n\cdot N^n)$ & \cite{roerdnik}\\
      $L_1$ & $O(1)$ &   $O(1)$ &$\Theta(n\cdot N^n)$& \cite{roerdnik}\\
      $L_p$  (exact pred.) & $O(\log{p})$ &
      $O(\log{p}\cdot\log{N})$ & $O(n\cdot
      N^n\cdot\log{p}\cdot\log{N})$& Lem.~\ref{lem:generic},  \cite{dgtal}\\
      $L_p$  (inexact pred.) & $O(1)$ &
      $O(\log{N})$ & $O(n\cdot
      N^n\cdot\log{N})$& Lem.~\ref{lem:generic},  \cite{dgtal}\\
      2D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
      &$O(\log^2{m}\cdot N^2)$& Theorem \ref{them}\\
  %    3D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
  %    &$O(\log^2{m}\cdot N^3)$& Theorem \ref{them2}\\
  %    nD Chamfer norm &  $O(n\cdot\log{f})$ &$O(n\cdot log^2{f}\cdot\log{N})$
  %    & $O(n^2\cdot N^n\cdot \log{f}\cdot\log{N})$& Corollary \ref{coro2}\\
      2D Neig. seq. norm  & $O(\log{f})$&
      $O(\log^2{f})$& $O(\log^2{}f\cdot N^2)$& Th.~\ref{them} and  \cite{DBLP:conf/dgci/NormandSE13}\\
      \hline
    \end{tabular}
  \end{center}
%\end{table}

\section{Preliminaries}
\label{sec:preliminaries}
%% Let us first define some elementary concepts and objects.
%% \begin{definition}[Metrics and Metric Space]
%%   \label{def:distance}
%%   A metric is a map $d$ from $E$ to a sub-group $F$ of $\R$ such that
%%   $\forall a,b,c\in E$,
%%   \begin{align}
%%     (\text{non-negative})\quad & d(a,b)\geq 0\\
%%     (\text{identity of indiscernibles}) \quad&  d(a,b)= 0
%%     \Leftrightarrow a=b\\
%%     (\text{symmetry})\quad &  d(a,b)=d(b,a)\\
%%     (\text{triangular inequality})\quad &   d(a,c) \leq d(a,b) + d(b,c)
%%   \end{align}
%% $(E, F, d)$ is called a metric space.
%% \end{definition}
%% If property $(4)$ is omitted, we are defining a distance function
%% instead of a metric.
\begin{definition}[Norm and metric induced by a norm]
  \label{def:norm}
  Given a vector space EV, a norm is a map $g$ from  $EV$ to a sub-group
  $F$ of $\R$ such that $\forall \vec{x},\vec{y}\in EV$,
  \begin{align}
    (\text{non-negative})\quad & g(\vec{x})\geq 0\\
    (\text{identity of indiscernibles}) \quad&  g(\vec{x})= 0
    \Leftrightarrow \vec{x}=\vec{0}\\
    (\text{triangular inequality})\quad &   g(\vec{x}+\vec{y}) \leq
    g(\vec{x}) + g(\vec{y})\\
    (\text{homogeneity})\quad &  \forall \lambda\in\R, \quad
    g(\lambda\cdot\vec{x}) = |\lambda|\cdot g(\cdot\vec{x})
  \end{align}
$d(a,b) \EqDef g(b-a)$ is the metric induced by the
  norm $g$.  $(E, F, d)$ is called a  metric space if $d:~E\rightarrow
  F$.
\end{definition}
Note that the above definition can be extended from vector spaces to
\emph{modules} on a commutative ring ($\Z^n$ being a module on $\Z$
but not a vector space) \cite{Thiel_hdr}. Path-based approaches
(chamfer masks, -weighted- neighborhood sequences...)  aim at defining
metrics induced by norms in metric spaces $(\Z^n, \Z , d)$.  Note that
(weighed, with $w_i\geq 0$) $L_p$ metrics
\begin{equation}
    d_{L_p} (a,b) = \left ( \sum_{k=1}^n w_k|a_k-b_k |^p \right )^{\frac{1}{p}}\;,
  \end{equation}
characterize metric spaces $(\Z^n, \R , d_{L_p})$. Rounding the \sloppy
distance function $(\Z^n,\Z, \lceil d_{L_p}\rceil)$ is a metric
space \cite{klette_book}.

@@

\cite{ROSEN_66,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13}

@@

In \cite{Hirata1996,Breu1995,Meijster2000,Maurer2003},

@@

consider Voronoi map approach as defined in \cite{Breu1995}: Let us

@@

In other words, \textsc{HiddenBy} returns true if and only if the
Voronoi cell of sites $a$ and $c$ \emph{hide} the Voronoi cell of $b$
along $S$.  For $L_1$, $L_2$ and $L_\infty$ metrics,
\textsc{Closest} and \textsc{HiddenBy} predicates can be computed in
$O(1)$ \cite{Breu1995,Hirata1996,Meijster2000}. Hence, Algorithm
\ref{alg:voro} is in $O(N_i)$ for dimension $i$, leading to an overall
computational time for the Voronoi map and distance transformation
problem in $\Theta(n\cdot N^n)$ (if we assume that $\forall i\in[1\ldots
  n], N_i=N$).

In \cite{Hirata1996} or \cite{maurer_pami}, authors discussed about
conditions on the metric $d$ to ensure that Algorithm \ref{alg:voro}
is correct. The key property can be informally describe as follows:
given two points $a,b\in\Z^n$ such that $a_i<b_i$ and a straight line
$l$ along the $i^{th}$ direction and if we denote by $v_l(a)$
(resp. $v_l(b)$) the intersection between the Voronoi cell of $a$
(resp. $b$) and l, then $v_l(a)$ and $v_l(b)$ are simply connected
Euclidean segments and $v_l(a)$ appears \emph{before} $v_l(b)$ on $l$
(so called \emph{monotonicity property} in \cite{Maurer2003} and is
related to \emph{quadrangle inequality} in \cite{Hirata1996}).  To sum
up these contributions, we have the following proposition:
\begin{proposition}[Metric conditions \cite{Hirata1996}]
\label{prop:voronoi-map-from-1}
  Let $d$ be a metric induced by a \textbf{norm whose unit ball is
    symmetric with respect to grid axes} and if distance comparison
  predicate is exact, Algorithm 1 is correct and returns a Voronoi map
  $\Pi_X$.
\end{proposition}
When implementing Algorithm 1, we say that the distance comparison
predicate is exact if we can compare two distances, \emph{e.g.}
\textsc{Closest} predicate, without error. Hence, Algorithm \ref{alg:voro}
provides exact Voronoi map computation for $L_p$ norms. In this case,
distance comparison predicate can be implemented error-free from
integer number comparison considering the $p$ power of the distance
function $\left (d_{L_p}\right)^p$.  Proposition
\ref{prop:voronoi-map-from-1} also implies that most chamfer norms and
neighborhood sequences based norms can also be considered in separable
Algorithm \ref{alg:voro} (see Fig.~\ref{fig:metrics}). We just need
algorithmic tools to efficiently implement both \textsc{Closest} and
\textsc{HiddenBy} predicates.
\begin{figure}[htbp]
 {\includegraphics[width=1.65cm]{image-ball-1}}
 {\includegraphics[width=1.65cm]{image-ball-1-4}}
 {\includegraphics[width=1.65cm]{image-ball-2}}
 {\includegraphics[width=1.65cm]{image-ball-4}}
 {\includegraphics[width=1.65cm]{image-ball-43-1}}
 {\includegraphics[width=1.65cm]{image-ball-chamf34}}
 {\includegraphics[width=1.65cm]{image-ball-chamf5711}}
  \caption{Balls for different metrics satisfying Proposition \ref{prop:voronoi-map-from-1}: \emph{(from
      left to right)} $L_1$, $L_{1.4}$, $L_2$, $L_4$, $L_{43.1}$,
    $\M_{3-4}$ and $\M_{5-7-11}$.}
  \label{fig:metrics}
\end{figure}
%
\vspace{-1cm}
\subsection{A First Generic Adapter}
\label{sec:first-gener-adapt}
We first detail the overall computational cost of Algorithm
\ref{alg:voro}:
\begin{lemma}
\label{lem:generic}
Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
symmetric unit ball. If $C$ denotes the computational cost of
\textsc{Closest} predicate and $H$ is the computational cost of the
\textsc{HiddenBy} predicate, then Algorithm \ref{alg:voro} is in
$O(n\cdot N^n\cdot (C+H))$.
\end{lemma}
From \cite{Breu1995,roerdnik}, $C=H=O(1)$ for $L_1,L_2$ and $L_\infty$
norms.  For prescribed norm $d$, we first define generic
Algorithms~\ref{alg:genclosest}, \ref{alg:genedge} and
\ref{alg:genhidden}: from some evaluations of $d$, \textsc{HiddenBy}
predicate is obtained by a dichotomic search on the 1D image span $S$
to localize the abscissa of Voronoi edges of sites $\{a,b\}$ and
$\{b,c\}$ (see Fig.~\ref{fig:algo}).

@@

In this section, we give some implementation details and experimental
results for chamfer norm Voronoi map in dimension 2.  First of all,
most algorithms presented here are available in the \textsc{DGtal}
library \cite{dgtal}. For $L_p$ metrics, we have implemented several
\textsc{Closest} and \textsc{HiddenBy} predicates: If $p=\{1,2\}$,
exact computations are proposed and all predicates are in $O(1)$ with
only integer number computations
\cite{Hirata1996,Maurer2003,roerdnik}; If $p\in \R$, $p\geq 1$, we
have approximated computations on real numbers (\emph{double}) and we
consider the Generic \textsc{HiddenBy} predicate in $O(\log N)$
(Alg.~\ref{alg:genhidden}). Since predicates are based on floating
point computations, numerical issues may occur. If $p\in \Z$, $p\geq
3$, we use exact integer number based computations of distances
storing sum of power $p$ quantities (which can be computed in
$O(\log{p})$ thanks to exponentiation by squaring). The
\textsc{HiddenBy} predicate is also based on
Algorithm~\ref{alg:genhidden}.
Beside these predicates for $L_p$ metrics, \textsc{DGtal} also
contains a generic metric adapter: if the user specifies a distance
function (taking two points and returning a value) corresponding to a
norm, generic \textsc{Closest} and \textsc{HiddenBy} predicates can be
automatically constructed. Please note that since all algorithms are
separable, generic framework provided in \textsc{DGtal} allow us to
have a free multi-thread implementation \cite{dcoeurjo_ChapDTWADGMM}.

>>>>

Since early works on digital geometry, distance transformation plays
an important role in many applications
\cite{Rosenfeld1966,Rosenfeld1968}. Given a finite input shape
$X\subset \Z^n$, the distance transformation labels each point in $X$
with the distance to its closest point in $\Z^n \setminus X$ (labeling
each point by the closest background point leads to Voronoi maps, see
below). Since such characterization is parametrized by a distance
function, many works have been done to address this distance
transformation problem with a trade-off between algorithmic
performances and the \emph{accuracy} of the digital distance function
with respect to the Euclidean one.  Hence, authors have considered
distances based on chamfer masks
\cite{Rosenfeld1968,borgefors,fouard:ivc:2005} or sequences of chamfer
masks \cite{Rosenfeld1966,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13};
the vector displacement based Euclidean distance
\cite{danielson,ragnemalm}; the Voronoi diagram based Euclidean
distance \cite{BreuEtAl95,Maurer2003} or the square of the Euclidean
distance \cite{Hirata,roerdnik}.  For Euclidean or $L_p$ metrics,
separable volumetric computations have demonstrated to be very
efficient: optimal $O(n\cdot N^n)$ time algorithms for shapes in $N^n$
domains, optimal multi-thread/GPU implementation\ldots (please refer
to \cite{dcoeurjo_ChapDTWADGMM} for a discussion). For path-based
metrics (chamfer mask, -weighted- neighborhood sequences,\ldots), two
main techniques exist to compute the distance transformation. The
first one considers a weighted graph formulation of the problem, and
shortest path in graphs using Dijkstra-like algorithm can be
considered. If $m$ denotes the size of the chamfer mask, computational
cost could be in $O(m\cdot N^n)$ using a cyclic bucket data structure
\cite{verwer_uniform}. Another approach consists in a raster scan of
the domain: first the chamfer mask is decomposed into two disjoint
sub-masks. Then the domain grid points are scanned in a given order
(consistent with the sub-masks construction) and a local minimum
computation is performed and then propagated
\cite{Rosenfeld1966,borgefors}. Scanning the domain twice for each
sub-masks leads to the distance transformation. Again, we end up with
a $O(m\cdot N^n)$ computational cost. This complexity is said to be
quadratic in $m$ and $N$ since for arbitrary large image $N$, even if
$m \ll N$, $m$ has to be in $O(N^n)$ to have accurate asymptotic DT.
Please note that Dijkstra's graph approach allows us to defined
constrained distance transformation (\emph{i.e.} geodesic metric),
both separable approaches and raster-scan for path-based metrics are
only dedicated to compact convex domains (usually hyper-rectangular)
distance transformation.

\textbf{Contributions} In this article, we first describe the generic
framework for separable distance transformation and conditions on the
metric to be consistent with this model. Then, we describe
subquadratic algorithms in dimension 2 to compute error-free distance
transformation and Voronoi map for chamfer norms and other path-based
metrics. Overall computational costs can be summarized as follows:
%\begin{table}
%  \caption{Computational cost summary for separable Voronoi map
%    computation on $N^n$ domains ($m$ being the size of the chamfer
%    norm and $f$ the number of row in a H-representation, see below).}
 % \label{tab:final}
  \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      Metric &\textsc{Closest}& \textsc{HiddenBy} & Sep. Voronoi Map & Reference\\
      \hline
      $L_2$ & $O(1)$ &  $O(1)$ & $\Theta(n\cdot N^n)$ & \cite{Hirata}\\
      $L_\infty$ & $O(1)$ & $O(1)$ &  $\Theta(n\cdot N^n)$ & \cite{roerdnik}\\
      $L_1$ & $O(1)$ &   $O(1)$ &$\Theta(n\cdot N^n)$& \cite{roerdnik}\\
      $L_p$  (exact pred.) & $O(\log{p})$ &
      $O(\log{p}\cdot\log{N})$ & $O(n\cdot
      N^n\cdot\log{p}\cdot\log{N})$& Lem.~\ref{lem:generic},  \cite{dgtal}\\
      $L_p$  (inexact pred.) & $O(1)$ &
      $O(\log{N})$ & $O(n\cdot
      N^n\cdot\log{N})$& Lem.~\ref{lem:generic},  \cite{dgtal}\\
      2D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
      &$O(\log^2{m}\cdot N^2)$& Theorem \ref{them}\\
  %    3D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
  %    &$O(\log^2{m}\cdot N^3)$& Theorem \ref{them2}\\
  %    nD Chamfer norm &  $O(n\cdot\log{f})$ &$O(n\cdot log^2{f}\cdot\log{N})$
  %    & $O(n^2\cdot N^n\cdot \log{f}\cdot\log{N})$& Corollary \ref{coro2}\\
      2D Neig. seq. norm  & $O(\log{f})$&
      $O(\log^2{f})$& $O(\log^2{}f\cdot N^2)$& Th.~\ref{them} and  \cite{DBLP:conf/dgci/NormandSE13}\\
      \hline
    \end{tabular}
  \end{center}
%\end{table}
\section{Preliminaries}
\label{sec:preliminaries}
%% Let us first define some elementary concepts and objects.
%% \begin{definition}[Metrics and Metric Space]
%%   \label{def:distance}
%%   A metric is a map $d$ from $E$ to a sub-group $F$ of $\R$ such that
%%   $\forall a,b,c\in E$,
%%   \begin{align}
%%     (\text{non-negative})\quad & d(a,b)\geq 0\\
%%     (\text{identity of indiscernibles}) \quad&  d(a,b)= 0
%%     \Leftrightarrow a=b\\
%%     (\text{symmetry})\quad &  d(a,b)=d(b,a)\\
%%     (\text{triangular inequality})\quad &   d(a,c) \leq d(a,b) + d(b,c)
%%   \end{align}
%% $(E, F, d)$ is called a metric space.
%% \end{definition}
%% If property $(4)$ is omitted, we are defining a distance function
%% instead of a metric.
\begin{definition}[Norm and metric induced by a norm]
  \label{def:norm}
  Given a vector space EV, a norm is a map $g$ from  $EV$ to a sub-group
  $F$ of $\R$ such that $\forall \vec{x},\vec{y}\in EV$,
  \begin{align}
    (\text{non-negative})\quad & g(\vec{x})\geq 0\\
    (\text{identity of indiscernibles}) \quad&  g(\vec{x})= 0
    \Leftrightarrow \vec{x}=\vec{0}\\
    (\text{triangular inequality})\quad &   g(\vec{x}+\vec{y}) \leq
    g(\vec{x}) + g(\vec{y})\\
    (\text{homogeneity})\quad &  \forall \lambda\in\R, \quad
    g(\lambda\cdot\vec{x}) = |\lambda|\cdot g(\cdot\vec{x})
  \end{align}
$d(a,b) \EqDef g(b-a)$ is the metric induced by the
  norm $g$.  $(E, F, d)$ is called a  metric space if $d:~E\rightarrow
  F$.
Note that the above definition can be extended from vector spaces to
\emph{modules} on a commutative ring ($\Z^n$ being a module on $\Z$
but not a vector space) \cite{Thiel_hdr}. Path-based approaches
(chamfer masks, -weighted- neighborhood sequences...)  aim at defining
metrics induced by norms in metric spaces $(\Z^n, \Z , d)$.  Note that
(weighed, with $w_i\geq 0$) $L_p$ metrics
    d_{L_p} (a,b) = \left ( \sum_{k=1}^n w_k|a_k-b_k |^p \right )^{\frac{1}{p}}\;,
characterize metric spaces $(\Z^n, \R , d_{L_p})$. Rounding the \sloppy
distance function $(\Z^n,\Z, \lceil d_{L_p}\rceil)$ is a metric
space \cite{klette_book}.

@@

\cite{Rosenfeld1966,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13}

@@

In \cite{Hirata,BreuEtAl95,roerdnik,Maurer2003},

@@

consider Voronoi map approach as defined in \cite{BreuEtAl95}: Let us

@@

In other words, \textsc{HiddenBy} returns true if and only if the
Voronoi cell of sites $a$ and $c$ \emph{hide} the Voronoi cell of $b$
along $S$.  For $L_1$, $L_2$ and $L_\infty$ metrics,
\textsc{Closest} and \textsc{HiddenBy} predicates can be computed in
$O(1)$ \cite{BreuEtAl95,Hirata,roerdnik}. Hence, Algorithm
\ref{alg:voro} is in $O(N_i)$ for dimension $i$, leading to an overall
computational time for the Voronoi map and distance transformation
problem in $\Theta(n\cdot N^n)$ (if we assume that $\forall i\in[1\ldots
  n], N_i=N$).

In \cite{Hirata} or \cite{Maurer2003}, authors discussed about
conditions on the metric $d$ to ensure that Algorithm \ref{alg:voro}
is correct. The key property can be informally describe as follows:
given two points $a,b\in\Z^n$ such that $a_i<b_i$ and a straight line
$l$ along the $i^{th}$ direction and if we denote by $v_l(a)$
(resp. $v_l(b)$) the intersection between the Voronoi cell of $a$
(resp. $b$) and l, then $v_l(a)$ and $v_l(b)$ are simply connected
Euclidean segments and $v_l(a)$ appears \emph{before} $v_l(b)$ on $l$
(so called \emph{monotonicity property} in \cite{Maurer2003} and is
related to \emph{quadrangle inequality} in \cite{Hirata}).  To sum
up these contributions, we have the following proposition:
\begin{proposition}[Metric conditions \cite{Hirata}]
\label{prop:voronoi-map-from-1}
  Let $d$ be a metric induced by a \textbf{norm whose unit ball is
    symmetric with respect to grid axes} and if distance comparison
  predicate is exact, Algorithm 1 is correct and returns a Voronoi map
  $\Pi_X$.
\end{proposition}
When implementing Algorithm 1, we say that the distance comparison
predicate is exact if we can compare two distances, \emph{e.g.}
\textsc{Closest} predicate, without error. Hence, Algorithm \ref{alg:voro}
provides exact Voronoi map computation for $L_p$ norms. In this case,
distance comparison predicate can be implemented error-free from
integer number comparison considering the $p$ power of the distance
function $\left (d_{L_p}\right)^p$.  Proposition
\ref{prop:voronoi-map-from-1} also implies that most chamfer norms and
neighborhood sequences based norms can also be considered in separable
Algorithm \ref{alg:voro} (see Fig.~\ref{fig:metrics}). We just need
algorithmic tools to efficiently implement both \textsc{Closest} and
\textsc{HiddenBy} predicates.
\begin{figure}[htbp]
 {\includegraphics[width=1.65cm]{image-ball-1}}
 {\includegraphics[width=1.65cm]{image-ball-1-4}}
 {\includegraphics[width=1.65cm]{image-ball-2}}
 {\includegraphics[width=1.65cm]{image-ball-4}}
 {\includegraphics[width=1.65cm]{image-ball-43-1}}
 {\includegraphics[width=1.65cm]{image-ball-chamf34}}
 {\includegraphics[width=1.65cm]{image-ball-chamf5711}}
  \caption{Balls for different metrics satisfying Proposition \ref{prop:voronoi-map-from-1}: \emph{(from
      left to right)} $L_1$, $L_{1.4}$, $L_2$, $L_4$, $L_{43.1}$,
    $\M_{3-4}$ and $\M_{5-7-11}$.}
  \label{fig:metrics}
\end{figure}
%
\vspace{-1cm}
\subsection{A First Generic Adapter}
\label{sec:first-gener-adapt}
We first detail the overall computational cost of Algorithm
\ref{alg:voro}:
\begin{lemma}
\label{lem:generic}
Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
symmetric unit ball. If $C$ denotes the computational cost of
\textsc{Closest} predicate and $H$ is the computational cost of the
\textsc{HiddenBy} predicate, then Algorithm \ref{alg:voro} is in
$O(n\cdot N^n\cdot (C+H))$.
\end{lemma}
From \cite{BreuEtAl95,roerdnik}, $C=H=O(1)$ for $L_1,L_2$ and $L_\infty$
norms.  For prescribed norm $d$, we first define generic
Algorithms~\ref{alg:genclosest}, \ref{alg:genedge} and
\ref{alg:genhidden}: from some evaluations of $d$, \textsc{HiddenBy}
predicate is obtained by a dichotomic search on the 1D image span $S$
to localize the abscissa of Voronoi edges of sites $\{a,b\}$ and
$\{b,c\}$ (see Fig.~\ref{fig:algo}).

@@

In this section, we give some implementation details and experimental
results for chamfer norm Voronoi map in dimension 2.  First of all,
most algorithms presented here are available in the \textsc{DGtal}
library \cite{dgtal}. For $L_p$ metrics, we have implemented several
\textsc{Closest} and \textsc{HiddenBy} predicates: If $p=\{1,2\}$,
exact computations are proposed and all predicates are in $O(1)$ with
only integer number computations
\cite{Hirata,Maurer2003,roerdnik}; If $p\in \R$, $p\geq 1$, we
have approximated computations on real numbers (\emph{double}) and we
consider the Generic \textsc{HiddenBy} predicate in $O(\log N)$
(Alg.~\ref{alg:genhidden}). Since predicates are based on floating
point computations, numerical issues may occur. If $p\in \Z$, $p\geq
3$, we use exact integer number based computations of distances
storing sum of power $p$ quantities (which can be computed in
$O(\log{p})$ thanks to exponentiation by squaring). The
\textsc{HiddenBy} predicate is also based on
Algorithm~\ref{alg:genhidden}.
Beside these predicates for $L_p$ metrics, \textsc{DGtal} also
contains a generic metric adapter: if the user specifies a distance
function (taking two points and returning a value) corresponding to a
norm, generic \textsc{Closest} and \textsc{HiddenBy} predicates can be
automatically constructed. Please note that since all algorithms are
separable, generic framework provided in \textsc{DGtal} allow us to
have a free multi-thread implementation \cite{dcoeurjo_ChapDTWADGMM}.
